Analysis_manuscript_mash_GxE

### 2020-11-21 R/Analysis_manuscript_mash_GxE.Rmd
Mashattan R scripts with allele frequency and subpop information

```{r}
m <- readRDS(paste0("../analysis/gwas/BLUPs/single_sites/Strong_Effects5000SNPs_greenup_2_phe_Gulf_and_Midwest_subpop_Gulf_Uhyp.rds"))

@@

Gulf
```{r}
inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf")
mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
mash_out
m <- read_rds(file.path(inputdir, mash_out[11]))
mash_plot_covar(m, saveoutput = TRUE)
get_U_by_mass(m)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_effects(m, n = 1, saveoutput = TRUE)
mash_plot_sig_by_condition(m, saveoutput = TRUE)
mash_plot_manhattan_by_condition(m, saveoutput = TRUE)
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[1])) # sign
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[2])) # sign & size
Gulf_greenup_gxe <- get_GxE(m)

>>>>

```{r}
m <- readRDS(paste0("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects5000SNPs_greenup_2_phe_Gulf_and_Midwest_subpop_Gulf_Uhyp.rds"))

@@

## Load m for Gulf
```{r}
inputdir <- file.path("../analysis/gwas/BLUPs/single_sites", "Gulf_and_Midwest")
mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
mash_out
m <- read_rds(file.path(inputdir, mash_out[13]))
m_meta <- read_rds("../analysis/Metadata_associated_with_alternate_alleles_snp_630g_PKLE_m_greenup_2_phe_Gulf_subpop_full.rds")
m_meta <- read_rds("../analysis/Metadata_associated_with_alternate_alleles_snp_363g_m_greenup_2_phe_Gulf_and_Midwest_subpop.rds")
```

Mashhattan where points are colored by which covariance matrix they load most strongly on to. Do different regions of the genome have distinct patterns, or are they all jumbled up?
```{r}
log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
    as.data.frame() %>%
    rownames_to_column(var = "value") %>%
    mutate(value = as.integer(.data$value)) %>%
    as_tibble() %>%
    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
    dplyr::rename(log10BayesFactor = .data$V1) %>%
    dplyr::select(-.data$value)

#' Return the name of the covariance matrix with the highest posterior weight
#'     for each effect
#'
#' @param m the mash result (from joint or 1by1 analysis)
#'
#' @return returns a df with Marker, Covariance_Matrix, and Grid_Value. I'm assuming that the grid value is an indication of like a svd decomp
#'
get_topcov = function(m){
  df <- switchgrassGWAS:::get_marker_df(m = m)
  covdf <- purrr::map_chr(df$value, ~ names(which(m$posterior_weights[.x,] ==
                                           max(m$posterior_weights[.x,])))) %>%
    enframe(name = "value", value = "Covariance_Grid") %>%
    separate("Covariance_Grid", into = c("Covariance_Matrix", "Grid_Value"),
             sep = "\\.", remove = FALSE) 
  df <- df %>% full_join(covdf, by = "value")
  return(df)
  }

df <- get_topcov(m)

  ggman_df <- df %>%
    left_join(m_meta) %>%
    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    arrange(.data$Chr, .data$Pos)

  log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))

  ggmanobject <- ggman_df %>%
    group_by(Covariance_Grid) %>% add_tally() %>%
    arrange(desc(n)) %>%
    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
    switchgrassGWAS::theme_oeco +
    geom_point(aes(color = .data$Covariance_Grid, fill = .data$Covariance_Grid,
                   shape = as.factor(.data$Chr))) +
    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
    scale_color_viridis_d(option = "D", end = 0.95) + scale_fill_viridis_d(option = "D", end = 0.95) +
    theme(strip.text = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          axis.text.x = element_blank(),
          legend.text = element_text(size = 8),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA)) +
    labs(x = "Chromosome", y = log10BF) +
    scale_x_continuous(expand = c(0.32, 0.32)) +
    scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
    geom_hline(yintercept = 2, linetype = 2)

ggman_df %>%
    group_by(Covariance_Grid) %>% add_tally() %>%
    arrange(desc(n)) %>%
    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
    switchgrassGWAS::theme_oeco +
    geom_point(aes(color = .data$Covariance_Grid, fill = .data$Covariance_Grid,
                   alpha = Alt_freq,
                   shape = as.factor(.data$Chr))) +
    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
    scale_color_viridis_d(option = "D", end = 0.95) + scale_fill_viridis_d(option = "D", end = 0.95) +
    theme(strip.text = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          axis.text.x = element_blank(),
          legend.text = element_text(size = 8),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA)) +
    labs(x = "Chromosome", y = log10BF) +
    scale_x_continuous(expand = c(0.32, 0.32)) +
    scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
    geom_hline(yintercept = 2, linetype = 2)

ggman_df %>%
  #arrange((Alt_Latitude)) %>%
    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
    switchgrassGWAS::theme_oeco +
    geom_point(aes(color = .data$Alt_Midwest_c,
                   #alpha = Alt_freq,
                   shape = as.factor(.data$Chr))) +
    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", 
               strip.position = "bottom") +
    scale_color_viridis_c(option = "B", end = 0.95) +
    #scale_fill_viridis_d(option = "D", end = 0.95) +
    theme(strip.text = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          axis.text.x = element_blank(),
          legend.text = element_text(size = 8),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA)) +
    labs(x = "Chromosome", y = log10BF) +
    scale_x_continuous(expand = c(0.32, 0.32)) +
    scale_shape_manual(values = rep(c(15, 16),9), guide = FALSE) +
    geom_hline(yintercept = 2, linetype = 2)

ggman_df
```

enrichment of sig SNPs for these different Alt_features.
Enrichment of covariance matrices for these different Alt_features
```{r}
library(lubridate)
metadata1 <- pvdiv_metadata %>%
  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
    mutate(Atlantic = ifelse(GENETIC_SUBPOP_KINSHIP == "Atlantic", 1, 0),
           Gulf = ifelse(GENETIC_SUBPOP_KINSHIP == "Gulf", 1, 0),
           Midwest = ifelse(GENETIC_SUBPOP_KINSHIP == "Midwest", 1, 0),
           Coastal = ifelse(ECOTYPE_NNET == "Coastal", 1, 0),
           Upland = ifelse(ECOTYPE_NNET == "Upland", 1, 0),
           Lowland = ifelse(ECOTYPE_NNET == "Lowland", 1, 0),
           COLL_YEAR = year(COLL_DATE))

ave630 <- metadata1 %>%
  summarise(All_Latitude = mean(LATITUDE, na.rm = TRUE),
            All_Longitude = mean(-LONGITUDE, na.rm = TRUE),
            All_Elevation = mean(ELEVATION, na.rm = TRUE),
            All_Year = mean(COLL_YEAR, na.rm = TRUE),
            All_Coastal_d = mean(Coastal, na.rm = TRUE),
            All_Coastal_c = mean(COASTAL_ASSIGNMENT, na.rm = TRUE),
            All_Lowland_d = mean(Lowland, na.rm = TRUE),
            All_Lowland_c = mean(LOWLAND_ASSIGNMENT, na.rm = TRUE),
            All_Upland_d = mean(Upland, na.rm = TRUE),
            All_Upland_c = mean(UPLAND_ASSIGNMENT, na.rm = TRUE),
            All_Atlantic_d = mean(Atlantic),
            All_Atlantic_c = mean(ATLANTIC_Q, na.rm = TRUE),
            All_Gulf_d = mean(Gulf, na.rm = TRUE),
            All_Gulf_c = mean(GULF_Q, na.rm = TRUE),
            All_Midwest_d = mean(Midwest, na.rm = TRUE),
            All_Midwest_c = mean(MIDWEST_DA, na.rm = TRUE))

ave363 <- metadata1 %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  summarise(All_Latitude = mean(LATITUDE, na.rm = TRUE),
            All_Longitude = mean(-LONGITUDE, na.rm = TRUE),
            All_Elevation = mean(ELEVATION, na.rm = TRUE),
            All_Year = mean(COLL_YEAR, na.rm = TRUE),
            All_Coastal_d = mean(Coastal, na.rm = TRUE),
            All_Coastal_c = mean(COASTAL_ASSIGNMENT, na.rm = TRUE),
            All_Lowland_d = mean(Lowland, na.rm = TRUE),
            All_Lowland_c = mean(LOWLAND_ASSIGNMENT, na.rm = TRUE),
            All_Upland_d = mean(Upland, na.rm = TRUE),
            All_Upland_c = mean(UPLAND_ASSIGNMENT, na.rm = TRUE),
            All_Atlantic_d = mean(Atlantic),
            All_Atlantic_c = mean(ATLANTIC_Q, na.rm = TRUE),
            All_Gulf_d = mean(Gulf, na.rm = TRUE),
            All_Gulf_c = mean(GULF_Q, na.rm = TRUE),
            All_Midwest_d = mean(Midwest, na.rm = TRUE),
            All_Midwest_c = mean(MIDWEST_DA, na.rm = TRUE))
```

```{r}
# --------

P1 <- ggman_df %>%
  mutate(Dev_Latitude = Alt_Latitude - ave363$All_Latitude,
         Dev_Longitude = Alt_Longitude - ave363$All_Longitude,
         Dev_Elevation = Alt_Elevation - ave363$All_Elevation,
         Dev_Year = Alt_Year - ave363$All_Year,
         Dev_Coastal = Alt_Coastal_d - ave363$All_Coastal_d,
         Dev_Lowland = Alt_Lowland_d - ave363$All_Lowland_d,
         Dev_Upland = Alt_Upland_d - ave363$All_Upland_d,
         Dev_Atlantic = Alt_Atlantic_d - ave363$All_Atlantic_d,
         Dev_Gulf = Alt_Gulf_d - ave363$All_Gulf_d,
         Dev_Midwest = Alt_Midwest_d - ave363$All_Midwest_d,) 
testvar <- "Dev_Midwest"
P1 <- P1 %>% rename(testvar = testvar)

lim_lower <- min(P1$testvar)
lim_upper <- max(P1$testvar) + max(P1$testvar) * 0.01

P1 %>%
  filter(log10BayesFactor>2) %>%
  ggplot(aes(x = testvar)) + 
  geom_histogram() + 
  geom_vline(xintercept = 0) +
  facet_wrap(~Covariance_Grid)

P1 %>%
  ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
  switchgrassGWAS::theme_oeco +
  geom_point(aes(color = .data$testvar,
                 #alpha = Alt_freq,
                 shape = as.factor(.data$Chr))) +
  facet_grid(.data$Covariance_Grid ~ .data$Chr, switch = "x", 
             scales = "free_x") +
  scale_color_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "white", "#27AD81FF", "#5DC863FF",
                                    "#FDE725FF"),
                      limits = c(lim_lower, lim_upper)) +
  theme(strip.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 8),
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
  labs(x = "Chromosome", y = log10BF) +
  scale_x_continuous(expand = c(0.32, 0.32)) +
  scale_shape_manual(values = rep(c(15, 16),10), guide = FALSE) +
  geom_hline(yintercept = 2, linetype = 2)
```

```{r}

mash_plot_covar(m, saveoutput = TRUE)
get_U_by_mass(m)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_effects(m, n = 1, saveoutput = TRUE)
mash_plot_sig_by_condition(m, saveoutput = TRUE)
mash_plot_manhattan_by_condition(m, saveoutput = TRUE)
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[1])) # sign
mash_plot_pairwise_sharing(effectRDS = file.path(inputdir, mash_out[2])) # sign & size
Gulf_greenup_gxe <- get_GxE(m)
